CRUSE to Safe Cycling in Ireland

### 2023-10-03 paper.qmd
Add route type section

CycleStreets' services can be used by anyone for free, but we commissioned a bespoke routing service for CRUSE for the following reasons:

- We needed to calculate hundreds-of-thousands of routes, which is against the terms of service of the free API.
- We needed to make changes to the routing profiles, including allowing routing on trunk roads, which are not permitted in the default routing profiles.
- We needed to be able to control the version of OSM data being used for the routing, enabling regular updates to the route networks as OSM data is updated.

>>>>

CycleStreets' services can be used by anyone for free, but we commissioned a bespoke routing service for CRUSE for the following reasons, to enable:

- Calculation of hundreds-of-thousands of routes, which is against the terms of service of the free API.
- Making changes to the routing profiles, including allowing routing on trunk roads, which are not permitted in the default routing profiles.
- Control over the version of OSM data being used for the routing, allowing regular updates to the route networks as OSM data is updated.

We calculated three network types for each disaggregated ('jittered') OD pair: 'Fastest', 'Balanced' and 'Quietest'.
As outlined on the CycleStreets' [website](https://www.cyclestreets.net/help/journey/howitworks/), the 'Fastest' routes minimise journey time, accounting for traffic lights and surface type.
The 'Quietest' route type minimises busy sections of road, allowing high 'diversion factors' from the fastest route on quieter but less direct ways (often avoiding roads and interactions with motor vehicles altogether where possible).
The 'Balanced' route type is a compromise between the two, minimising journey time while avoiding the busiest roads.
Allowing users to switch between these three route types enables them to consider the trade-offs between directness and 'cycle friendliness' when planning new infrastructure, as shown in @fig-route-types.
